---
title: "STAT37810 Assignment1"
author: "Yanfei Zhou (UCID:12207462)"
output:
  word_document: default
  pdf_document: default
  html_document: default
Date: 11/10/2018
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 4.1.1
### 2a)
```{r}
Fibonacci <- numeric(30) 
Fibonacci[1] <- Fibonacci[2] <- 1  #assign first two element sin Fibonacci as 1,
for (i in 3:30){
Fibonacci[i] <- Fibonacci[i - 2] + Fibonacci[i - 1]
}
Fibonacci
Fib_ratio <-numeric(29)
for ( i in 2: 30){
  Fib_ratio[i] = Fibonacci[i]/Fibonacci[i-1]
}
Fib_ratio
```
The sequence appears to be converging to 1.618034


### 2 b)
```{r}
Golden_ratio = (1+sqrt(5))/2
Golden_ratio
```
The Fibonacci sequence converges to this ratio:

Let $\psi=\frac{1+\sqrt{5}}{2}$, notice $\psi=1+\frac{1}{\psi}$. Let $\big\{F_{n}\big\}$ with n=1,2,3,... be Fibonacci sequence. Let $R_{n}=\frac{F_{n+1}}{F_{n}}$, to prove: $$\lim_{n\to\infty} R_{n}=\psi=\frac{1+\sqrt{5}}{2}$$
\[
\begin{eqnarray}
R_{n} & = &\frac{F_{n+1}}{F_{n}}\\
& = &\frac{F_{n}+F_{n-1}}{F_{n}}\\
& = & 1 + \frac{1}{R_{n-1}}\\
|R_{n}-\psi| &= &|1+\frac{1}{R_{n-1}}-1-\frac{1}{\psi}|\\
& = & |\frac{\psi - R_{n-1}}{R_{n-1}\psi}|\\
& = &\frac{1}{\psi}|\frac{\psi-R_{n-1}}{R_{n-1}}|\\
& \leq & \frac{1}{\psi}|\psi-R_{n-1}| \; \;\;\;\:\;\;\;(as\;R_{n}>1)\\
& = & \frac{1}{\psi}|R_{n-1}-\psi|\\
& \leq & (\frac{1}{\psi})^{n}|R_{1}-\psi|\\
\lim_{n\to\infty}|R_{n}-\psi| &\leq&\lim_{n\to\infty}(\frac{1}{\psi})^{n}|R_{1}-\psi|\\
& \rightarrow &0 \;\;\;\;(as \; \frac{1}{\psi}<1)\\
\end{eqnarray}
\]
Hence we have $$\lim_{n\to\infty}|R_{n}| = \psi$$

#Question 4.1.1.

### 3 a)
answer = 15
```{r}
answer <- 0
for (j in 1:5) answer <- answer + j
answer
```
### 3 b)
answer = 1 2 3 4 5
```{r}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
answer
```
### 3 c) 
answer = 0 1 2 3 4 5
```{r}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
answer
```
### 3 d) 
answer = 120
```{r}
answer <- 1
for (j in 1:5) answer <- answer * j
answer
```
### 3 e)
3 21 23 6 11 15 12 22 30 24 13 29 17 26 27  3
```{r}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
answer
```
the sequence 3 21 23 6 11 15 12 22 30 24 13 29 17 26 27 will repeat when second 3 appears.

#Question 4.1.2 
### 4
```{r}
GIC<-function(P,t){
  if (t>3){
    return (P*((1 + 0.05)^t - 1))
  } else {
    return (P*((1 + 0.04)^t - 1))
  }}
GIC(100,3)
```

### 5
```{r}
Mortgage <-function(n,P,open){
  if (open==TRUE){
    i = 0.005
  } else{
    i=0.004
  }
  return (P*i/(1-(1+i)^(-n)))
}
Mortgage(3,100,FALSE)
```
#Question 4.1.3
### 2
```{r}
i=1
Fibonacci <- c(1, 1) 
while (i<30) {
  Fibonacci <- c(Fibonacci, Fibonacci[i-1]+Fibonacci[i]) 
  i = i+1
}
Fibonacci
```

### 4
```{r}
i <- 0.006
b = 1
a = 0
while (abs(b - a)>=0.000001) {
  b <- i
  i <- (1 - (1 + i)^(-20)) / 19
  a <- i 
}
#trying other starting values
i
i <- 0.005
b = 1
a = 0
while (abs(b - a)>=0.000001) {
  b <- i
  i <- (1 - (1 + i)^(-20)) / 19
  a <- i 
}
i
```
When using another starting guess, the point estimation becomes slightly different.

### 5 
```{r}
i <- 0.006
b = 1
a = 0
n = 0
while (abs(b - a)>=0.000001) {
  b <- i
  i <- (1 - (1 + i)^(-20)) / 19
  a <- i 
  n = n+1}
n
```


# Question 4.1.5
### 2 a)
```{r}
Eratosthenes<- function(n) {
  #find all the primes between 2 to n
  if (n>=2){
    sieve <- seq(2,n) #create a consecutive integer sequence from 2 up to n
    primes <- c() #create an empty list to store primes later
    while(length(sieve)>0){
      p<-sieve[1] #store the first element remaining in sieve as p
      primes<-c(primes,p) #add p to prime list
      sieve<-sieve[(sieve%%p)!=0] #remove all the elements in remaining sieve sequence that can be divided by p
    }
    return (primes) #return prime list 
  } else{
    stop("Input value of n should be at least 2.") }
}
Eratosthenes(11)
```

### 2 b)
```{r}
Eratosthenes<- function(n) {
  if (n>=2){
    sieve <- seq(2,n)
    primes <- c()
    while(length(sieve)>0){
      p<-sieve[1]
      if (p>=sqrt(n)){
        print(c("sieve: ",sieve))
        break} 
      primes<-c(primes,p)
      sieve<-sieve[(sieve%%p)!=0]
    }
    return(primes)
  } else{
    stop("Input value of n should be at least 2.") }
}
Eratosthenes(11)
```
Hence once $p \geq \sqrt{n}$, all the remaining entries in sieve are primes.


### 2 c)
```{r}
Eratosthenes<- function(n) {
  if (n>=2){
    sieve <- seq(2,n)
    primes <- c()
    while(length(sieve)>0){
      p<-sieve[1]
      if(p>=sqrt(n)){
        primes<-c(primes,sieve)
        break}
      primes<-c(primes,p)
      sieve<-sieve[(sieve%%p)!=0]
    }
    return (primes)
  } else{
    stop("Input value of n should be at least 2.") }
}
Eratosthenes(11)
```



# Question 4.2.1. 
### 2 a)
```{r}
compound.interest <- function(P, n, ir){
  return (P*((1+ir)^n))}
compound.interest(100,3,0.04)
```

### 2 b)
```{r}
compound.interest(1000,30,0.01)
```

He will have $1347.849 in the bank at the end of 30 months.

# Question 4.2.1. 
### 3
```{r}
bisection <- function(f, a, b, nmax=100, tol=1e-8){
  # check if f(a) and f(b) has opposite signs
  if(sign(f(a))==sign(f(b))) {
    stop('Oops! f(a) and f(b) have same signs. Need to change a or b to have f(a), f(b) opposite signs!')
    }
  n = 1
  while(n < nmax){
    # find midpoint
    c <- (a+b)/2 
    # if c is the root or the midpoint is below tolerance, output c
    if ((f(c)==0) | (b - a) / 2 < tol) return (c) 
    #check the signs of f(a) with f(c), and f(c) with f(b), reassign a or b accordingly as the value of c to be used in the next iteration.
    else if (sign(f(a))!=sign(f(c))) b <- c
    else a <- c
    n = n+1
  }
  #when root is not found after nmax iteration
  print("method failed, try more iterations or check function!")
}
f1 <- function(x) {
  x^3 - 2 * x - 5
}
bisection(f1,a=-5,b=3)
```


# Question 4.4.1. (1)
### 1
```{r}
mergesort <- function (x, descending) {
  len <-length(x)
  if (len < 2) result <- x 
  else {
    y <- x[1:(len / 2)]
    z <- x[(len / 2 + 1):len]
    y <- mergesort(y, descending)
    z <- mergesort(z, descending)
    result <- c()
    if (descending==FALSE){
      while (min(length(y), length(z)) > 0) {
        if (y[1] < z[1]) {
          result <- c(result, y[1])
          y <- y[-1]
          } else {
            result <- c(result, z[1])
            z <- z[-1]}}
      if (length(y) > 0) result <- c(result, y)
      else result <- c(result, z)
      return (result)
    } else if (descending==TRUE){
      while (min(length(y), length(z)) > 0) {
        if (y[1] > z[1]) {
          result <- c(result, y[1])
          y <- y[-1]
        } else {
          result <- c(result, z[1])
          z <- z[-1]}}
      if (length(y) > 0) result <- c(result,y)
      else result <- c(result,z)
      return (result)
    }
  }
}
mergesort(c(2,6,1,7,5,22),descending = TRUE)
mergesort(c(2,6,1,7,3,22),descending = FALSE)
```



### 2 a)
```{r}
newton<-function(f,g, tol=1e-8,nmax=100,x0,y0){ #define function with argument: function f, function g, minimum tolerance, maximum number of iterations nmax, set initial values x0, y0
  h = 1e-8
  i = 1
  while( i <= nmax){
    #find x1, y1 from the definition of derivatives and rules of Newton method
    df.dx = (f(x0+h,y0)-f(x0,y0))/h
    df.dy = (f(x0,y0+h)-f(x0,y0))/h
    dg.dx = (g(x0+h,y0)-g(x0,y0))/h
    dg.dy = (g(x0,y0+h)-g(x0,y0))/h
    x1 = x0 - (dg.dy*f(x0,y0) - df.dy*g(x0,y0))/(df.dx*dg.dy - df.dy*dg.dx)
    y1 = y0 - (df.dx*g(x0,y0) - dg.dx*f(x0,y0))/(df.dx*dg.dy - df.dy*dg.dx)
    i = i+1
    #if the tolerance between two iteration is smaller than threshold we set, then break and output x1, y1
    if (abs(x1-x0)<tol && abs(y1-y0)<tol) break
    #if the tolerance is not small enough, continue to next iteration
    x0 = x1
    y0 = y1
  }
  return(c(x1,y1))
}
```

### 2 b)
```{r}
f<-function(x,y) x^2+2*y^2-2
g<-function(x,y) x+y
newton(f=f,g=g,x0=1,y0=1)
newton(f=f,g=g,x0=-1,y0=-1)
```
Analytical solutions:
\[
\left\{
\begin{array}{ll}
x+y = 0\\
x^{2}+2y^{2}-2 = 0\\
                \end{array}
              \right.
\]
From the first equation we have $x=-y$, substitute into second equation, we get $y^{2}+2y^{2}-2=0$ , which gives  $y=\sqrt{\frac{2}{3}}$, one root coincides with the numerical solution we found above.


# Chapter 4 
### 1
```{r}
directpoly<-function(c,x){
  P=0
  for (i in 1:length(c)){
    P = P + c[i]*(x**(i-1))
  }
  return(P)
}
directpoly(c(1,5,2,6,2),c(1,2))
```

### 2
```{r}
hornerpoly<-function(c,x){
  a<-matrix(c,0,nrow=length(c),ncol=length(x))
  for (i in (length(c)-1):1){
    a[i,]=a[i+1,]*x+c[i]
  }
  return(a[1,])
}
hornerpoly(c(1,5,2,6,2),c(1,2))
```

### 3 a)
```{r}
system.time(directpoly(c=c(1,-2,2,3,4,6,7),x=seq(-10,10,length=5000000)))
system.time(hornerpoly(c=c(1,-2,2,3,4,6,7),x=seq(-10,10,length=5000000)))
```

Hornerpoly is more efficient when the number of polynomial coefficients is large.

### 3 b)
```{r}
system.time(directpoly(c=c(-3,17,2),x=seq(-10,10,length=5000000)))
system.time(hornerpoly(c=c(-3,17,2),x=seq(-10,10,length=5000000)))
```

When the number of polynomial coefficient is small, directpoly turns to be more efficient.









